---
title: "Sample data analysis using R report"
author: "Martín Lotto Batista"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, 
                      warning=FALSE, kableExtra.auto_format = FALSE)
```

```{r packages&data}
pacman::p_load(tidyverse, lubridate, sf, ggspatial, spdep, 
               kableExtra, cowplot, MASS)

data <- read.csv("data/clean_data.csv")
pop <- read.csv("data/population_data.csv")
study.area <- st_read("data/maps/NUTS3_germany_4326.shp", quiet=TRUE)
```

<style>
  body {text-align: justify}
</style>

```{r summary statistics, results='hide'}
# Total number of cases in dataset
sum(data$counts)

# Overall annual incidence rate
(sum(data$counts)/sum(pop$pop))*100000

# Annual incidence rate per district
map.data <- data %>% 
  group_by(nuts3, age_cat, sex) %>% 
  summarise(counts=sum(counts), .groups="drop") %>%
  left_join(pop) %>% 
  group_by(nuts1, nuts3, NUTS3_ID) %>% 
  summarise(counts=sum(counts),
            pop=sum(pop), .groups="drop") %>%
  # Compute incidence per 100,000 inhabitants
  mutate(incidence=(counts/pop)*100000) 

map.data %>% 
  # Order dataset in decreasing order
  arrange(desc(incidence)) %>% 
  # Remove districts with few counts as incidence is inflated in these contexts
  filter(counts > 5) %>% 
  head(n=5)
```

```{r figure1, out.width= "50%", out.extra='style="float:right; padding:10px"'}
# Create categories for incidence based on case distribution
map.data <- map.data %>% 
  mutate(incidence_cat=cut(incidence, 
                           breaks = seq(0, 150, length.out=6),
                           include.lowest = TRUE))

pal <- c("#f1eef6", "#d7b5d8", "#df65b0", "#dd1c77", "#980043")

pt1 <- study.area %>%
  # Add regrouped data
  left_join(map.data) %>% 
  ggplot() +
  # Plot grouped incidence
  geom_sf(aes(fill=incidence_cat), color="#d4b9de") +
  # Define palette (using ggthemes)
  scale_fill_manual(values=pal, 
                    na.translate=FALSE) +
  labs(fill="Cases per 100,00 inhabitants",
       title = "Figure 1") +
  # Add a layer for Federal states borders by regrouping the intial map
  geom_sf(data=study.area %>% 
            group_by(nuts1) %>% 
            summarise(.groups="drop"), col="black", fill=NA) +
  # Set aesthetic elements
  theme_void() +
  theme(legend.text=element_text(size=9),
        legend.title=element_text(face="bold", size=9),
        plot.title=element_text(size=12),
        plot.margin=unit(c(0,0,0.5,0), "cm")) +
  annotation_north_arrow(location="tl",
                                    which_north="true", 
                                    height=unit(0.85, "cm"), 
                                    width=unit(0.75, "cm")) +
  annotation_scale(location="bl", pad_y=unit(0, "cm"), line_width=0.5)

pt1
```

#### Introduction

This report utilizes data from the German surveillance system, which is publicly 
available on the online platform Survstat@RKI maintained by the [Robert
Koch Institute](https://survstat.rki.de/). The data used in this analysis includes 
weekly case counts of Lyme disease reported in 2021, stratified
by NUTS-3 district, age group and binary sex classification. Population size data 
for the corresponding year were extracted from the German Statistics Office 
[(Destatis)](https://www.destatis.de/DE/Home/_inhalt.html). For the purpose of this 
report, cases without spatial allocation or with unknown 
age or sex designation were excluded. In this document, a brief description of the
data will be provided using summary statistics, classic visualization techniques, 
and a basic exploratory analysis.

#### Lyme disease cases in Germany in 2021

In 2021, a total of 10,841 Lyme disease cases were reported to the German surveillance
system. The overall average annual incidence rate was 13.02 cases per 100,000 inhabitants. 
The districts with the highest annual incidence rates per 100,000 inhabitants were 
Straubing-Bogen, Pasau, Regen and Freyung-Grafenay in Bavaria, as well as Bandenburg city 
located in the state of Brandenburg (Figure 1).

Age showed a bimodal distribution with the first peak among children aged up to 10 years
old and the second peak among adults between 40 and 75 years old. Regarding sex 
distribution, the number of cases was higher in females than in males (Table 1). 
The majority of the Lyme disease cases occurred between late Spring (June) and 
early Autumn (October) (Figure 3). 

```{r table1}
# Age group and sex distribution
data %>% 
  group_by(sex, age_cat) %>% 
  summarise(counts=sum(counts), .groups="drop") %>%
  pivot_wider(names_from=sex, values_from=counts) %>% 
  mutate(across(2:3, ~paste0(.x, " (",round((.x/sum(.x))*100,2), "%)"))) %>% 
  rename(`Age group`=age_cat,
         Females=female,
         Males=male) %>% 
  kbl(caption="Table 1") %>%
  kable_classic(full_width=FALSE, html_font="Arial", position="center")
```

To assess the general spatial structure in the data, the Global Moran's I statistic 
was used. The results indicate that the spatial distribution of incidence among 
districts is random (Moran's I: -0.009, p-value: 0.522).

```{r figure2, fig.width=6, fig.height=4, fig.align='center'}
# Seasonality
pt2 <- data %>% 
  mutate(month=month(date, label=TRUE)) %>% 
  group_by(nuts3, month) %>% 
  summarise(counts=sum(counts), .groups="drop") %>%
  ggplot() +
  geom_boxplot(aes(x=month, y=counts)) +
  ylim(c(0,50)) +
  labs(x="", 
       y="Cases per 100,000 inhabitants",
       title="Figure 2") +
  theme_classic() +
  theme(axis.text = element_text(angle=45, hjust=1, size=12, face="bold"),
        axis.title.y=element_text(size=15),
        plot.title=element_text(size=12))

pt2
```

```{r MoransI, results='hide'}
# Compute Moran's I
# Neighbourhood matrix
nb <- poly2nb(study.area %>% 
                filter(nuts3 %in% unique(data$nuts3)), queen=T) 
# Queen to consider at least one vertex contact as neighbours

# Weights to neighbors
lw <- nb2listw(nb, style="W", zero.policy=TRUE)

# Hypothesis test where H0: not-randomly distributed
MC <- moran.mc(map.data$incidence, lw, alternative="greater", nsim=999)
MC  # MI: -0.009 and p-val: 0.522
```

```{r Poisson, results='hide'}
data.model <- data %>% 
  group_by(nuts3, sex, age_cat) %>% 
  summarise(counts=sum(counts), .groups="drop") %>% 
  left_join(pop) %>% 
  mutate(nuts3=as.numeric(factor(nuts3)))

# Check whether variance is similar to the mean
var(data.model$counts) # 64.7
mean(data.model$counts) # 4.7
# This indicates that there is overdispersion, hence I will use a negative binomial distribution

m <- glm.nb(counts ~ factor(sex) + factor(age_cat) + offset(log(pop)),
             data=data.model)

CI <- data.frame(estimate=m[["coefficients"]],
                 lower=confint(m)[,1],
                 upper=confint(m)[,2]) %>% 
  rownames_to_column("var") %>% 
  filter(!grepl("Intercept", var)) %>% 
  mutate(across(2:4, ~exp(.x)),
         var=str_replace_all(var, c("factor\\(sex\\)male"="Male",
                                    "factor\\(age_cat\\)"="Age group ")))
```

To evaluate the association between age groups, sex and the occurrence of Lyme
disease, a generalized linear model (GLM) was employed. Given the overdispersion
in the data, the response was assumed to follow a negative binomial distribution. 
Based on the results, after accounting for the role of age, the risk of Lyme disease
was higher in females than in males, with a relative risk (RR) for males of 0.9 
(CI: 0.84-0.96). Likewise, after considering the effect of sex, the highest risk 
was observed in the 60-75 age group, with a RR of 1.14 (CI: 1.01-1.29), while the 
lowest risk in the 10-20 and 20-30 age groups, with RRs of 0.56 (CI: 0.48-0.66) 
and 0.59 (CI: 0.50-0.69), respectively (Figure 3).

```{r figure3, fig.width=6, fig.height=4, fig.align='center'}
pt3 <- CI %>% 
  ggplot() +
  geom_pointrange(aes(x=var, y=estimate, ymin=lower, ymax=upper)) +
  geom_hline(yintercept=1, linetype="dashed", col="grey80") +
  labs(x="", 
       y="Relative Risk (RR)",
       title="Figure 3") +
  coord_flip() +
  theme_classic() +
  theme(axis.text.x=element_text(angle=45, hjust=1),
        axis.text=element_text(size=9),
        plot.title=element_text(size=9))

pt3
```

#### Conclusion
All the Federal States in Germany with ongoing Lyme disease surveillance reported 
cases to the national surveillance system. The highest incidence was found in 
Bavaria and Brandenburg states, and there were signs of a random spatial distribution 
of cases. The risk of disease was higher in females than in males, as well as in 
adults between 60 and 75 years old.

#### End